---
title: "Data Visulizations using ggplot"
output: html_document
---

```{r ,message=FALSE,warning=FALSE}
#Loading Packages
library(tidyverse)
```


#A)Basic Visulizations

##Note-To build a ggplot, we will use the following basic template that can be used for different types of plots:

```{r}
#ggplot(data = <DATA>, mapping = aes(<MAPPINGS>)) +  <GEOM_FUNCTION>()

```

##Use the ggplot() function and bind the plot to a specific data frame using the data argument

```{r}
#ggplot(data = surveys_complete)
```

###Note-define a mapping (using the aesthetic (aes) function), by selecting the variables to be plotted and specifying how to present them in the graph, e.g. as x/y positions or characteristics such as size, shape, color, etc.

```{r}
#ggplot(data = mtcars, mapping = aes(x = mpg, y = cyl))
```

###add 'geoms' - graphical representations of the data in the plot (points, lines, bars). ggplot2 offers many different geoms; we will use some common ones today, including:

###* `geom_point()` for scatter plots, dot plots, etc.
###* `geom_boxplot()` for, well, boxplots!
###* `geom_line()` for trend lines, time series, etc.  

###To add a geom to the plot use the + operator. Because we have two continuous variables, let's use  geom_point() first:

```{r}
#Scatter Plot
ggplot(data = airquality, mapping = aes(x = airquality$Day, y =Temp )) +
  geom_point()
```

##Assign a plot to a variable
```{r}
# Assign plot to a variable
aq_plot <- ggplot(data = airquality, mapping = aes(x = airquality$Day, y =Temp )) +
  geom_point()

# Draw the plot
aq_plot + 
    geom_point()
```

#Line Graph

```{r,,warning=FALSE}
ggplot(data = airquality, mapping = aes(x =airquality$Ozone, y =airquality$Temp )) +
  geom_line()
```

#Boxplot

```{r,warning=FALSE}

ggplot(data = airquality, mapping = aes(x =airquality$Ozone, y =airquality$Temp )) +
  geom_boxplot()

```

#Adding Labels and Titles
```{r,warning=FALSE}
ggplot(data = airquality, mapping = aes(x =airquality$Ozone, y =airquality$Temp )) +
  geom_line()+
  xlab("Ozone")+
  ylab("Temperature")+
  ggtitle("Ozone vs Temp")


```

#Level 2 Visulizations
```{r}
library(ggplot2)
theme_set(theme_classic())

# Plot
g <- ggplot(mpg, aes(cty))
g + geom_density(aes(fill=factor(cyl)), alpha=0.8) + 
    labs(title="Density plot", 
         subtitle="City Mileage Grouped by Number of cylinders",
         caption="Source: mpg",
         x="City Mileage",
         fill="# Cylinders")
```

#Violin Graphs
```{r}
theme_set(theme_bw())

# plot
g <- ggplot(mpg, aes(class, cty))
g + geom_violin() + 
  labs(title="Violin plot", 
       subtitle="City Mileage vs Class of vehicle",
       caption="Source: mpg",
       x="Class of Vehicle",
       y="City Mileage")
```

#Lollipop Charts

```{r}
#creating the dataset
cty_mpg <- aggregate(mpg$cty, by=list(mpg$manufacturer), FUN=mean)  # aggregate
colnames(cty_mpg) <- c("make", "mileage")  # change column names
cty_mpg <- cty_mpg[order(cty_mpg$mileage), ]  # sort
cty_mpg$make <- factor(cty_mpg$make, levels = cty_mpg$make)  # to retain the order in plot.
head(cty_mpg, 4)

# Plot
ggplot(cty_mpg, aes(x=make, y=mileage)) + 
  geom_point(size=3) + 
  geom_segment(aes(x=make, 
                   xend=make, 
                   y=0, 
                   yend=mileage)) + 
  labs(title="Lollipop Chart", 
       subtitle="Make Vs Avg. Mileage", 
       caption="source: mpg") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6))
```

#Advanced Visulizations

Key features of gganimate:

transitions: you want your data to change
views: you want your viewpoint to change
shadows: you want the animation to have memory
```{r}
library(gganimate)
ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_boxplot() + 
  # Here comes the gganimate code
  transition_states(
    gear,
    transition_length = 2,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('sine-in-out')
```

```{r}
library(gapminder)

ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear')
```

#transition_time(). The transition length between the states will be set to correspond to the actual time difference between them.

#facet_wrap-creating facet by countries.
#ease_aes-smooth transition between points.
